---
title: "Lab Assignment 1"
author: "Fill in your Name"
date: "Fill in the date"
output:
  pdf_document: default
  html_document: default
---

# Likelihood methods for the Poisson distribution

* Execute each chunk of code to ensure that your code works properly.  
* Then HTML knit the entire document.
* If you cannot compile all of your code without errors before the end of the class, 
comment out the chunks that are not working.
* Open the HTML document in a browser.
* Save the HTML file as a pdf file. 
* Submit your pdf assignment.

## 1.  First generate 1 observations from the Poisson(lambda=5) distribution and print the value.


```{r data}
set.seed(54321)   #setting the random number generator seed SHOULD always be done before simulating


```

## 2.  Compute the Log-likelihood for a vector sequence of lambda values from 2 to 7 in steps of .05 .


```{r LogLik}

```


## 3. Plot the Log-Likelihood function values from your answer in 2 versus lambda. Axes must be labelled and the plot must have a title.

```{r plotLog}

```


## 4.  Generate n=10 observations from the same distribution and compute the sample mean of the 10 observations.  Plot the log-likelihood function given the n=10 observations.  Axes must be labelled and the plot must have a title.


```{r plotLog10}
set.seed(54321)

```


